Mining Scenario/Transaction Pairs in HW/SW Co-verification Using Symbolic Execution

### 2017-04-07 paper.tex
fixes

%
%===============================================================================
\section{Introduction}
%===============================================================================
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                    Pre-RTL versus Post-RTL
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Hardware/software co-verification is broadly performed at two different
phases in the design cycle: pre-Register Transfer Level (pre-RTL) and 
post-RTL.  In the pre-RTL phase,
engineers write high-level software (SW) models for Intellectual Property (), 
which are commonly known as {\em Transaction Level Model}s
(TLM)~\cite{codes14}.  TLMs are typically implemented using the SystemC TLM
library~\cite{tlm} or as plain C programs.  The goal of a TLM is to capture
enough functionality of the hardware (HW) such that it is possible to execute the
SW (drivers or firmware (FW)) against it.  Consequently, the TLMs are
incomplete, as they are designed to contain only those behaviors that are
observable by the SW (the drivers or FW).  In particular, TLMs
are often not cycle-accurate.  These models are used to verify high-level
functionality of the system well before the RTL is
available~\cite{codes14,codes15}. Co-verification of the TLM and the
SW is typically performed by means of testing~\cite{hvc,codes14}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                    Challenges in Co-verification 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The verification of a FW given in C/C++ together with an RTL given in 
Verilog/VHDL is a challenging task due to a
few reasons -- 1) the timing and semantic mismatch between the synchronous
clock-driven HW model and asynchronous event-driven SW model. For example, the
FW running on a processor could be much faster than the HW
model it interacts with, 2) lack of support for property specification of 
co-designs that can precisely capture the design intent of the 
communication between FW and HW, 3) large
number of interleavings between HW and FW threads, and 4) lack 
of automated formal tools that support co-designs implemented in C/C++ and 
Verilog/VHDL.   In this paper, we address some of these challenges to apply 
formal techniques in hardware/software co-verification.

@@

\textbf{Step 2: Sequentialization based on Transaction Interaction}
%
Concurrency is at the heart of co-verification.  Testing concurrent 
threads requires exploring all possible interleavings between a HW 
and FW thread.  The number of interleavings could be potentially 
exponential.  However,~\cite{codes14,hvc} showed that the 
interaction between a FW thread and HW thread follows a 
producer-consumer relationship. That is, a FW 
thread is mostly independent of the HW thread they interact with~\cite{hvc,codes14}.  
Specifically, a FW thread is only responsible for configuring the 
HW transactions by writing to memory mapped registers or polling 
the interrupt status register for data transmission or receiving 
incoming data packets.

@@

Figure~\ref{fig:interleaving1} illustrates the producer-consumer relationship 
in UART . The FW CFG is shown in the left and the 
HW RTL code snippet is shown in the right. We abuse notations and 
represent $\mathcal{HT}$ in the form of RTL and $\mathcal{FT}$ as CFG
for our illustration purpose. Here, $\mathcal{HT}$ is the producer of interrupt 
status signal, $irq\_o$, which is read by $\mathcal{FT}$ in state $s_1$. 
The signal $irq\_o$ is  triggered when transmitter buffer is 
{\em EMPTY} or receiver buffer is {\em FULL}. 
$\mathcal{HT}$ also produces interrupt status by writing to the 
memory mapped register, $UART\_IS$, which is consumed by 
$\mathcal{FT}$ in state $s_2$.  In another interaction, $\mathcal{HT}$
sends a data-packet to $UART\_TR$ which is consumed by $\mathcal{FT}$ in state $s_7$.
Similarly, $\mathcal{FT}$ is the producer of $data$, $(outb(data,UART\_TR))$, 
in state $s_6$ which is consumed by $\mathcal{HT}$.  
%
\begin{figure}[t]
\scalebox{.55}{\import{figures/}{interleaving.pspdftex}}
\caption{Producer Consumer Interactions Pattern
\label{fig:interleaving1}}
\end{figure}
%

@@

\textbf{Step 3: Co-verification Model}
%
An unified co-verification model is constructed through appropriate composition of 
FW and its interacting HW transaction.  This is done by instrumenting the 
software netlist model of HW RTL inside the FW transaction at every invocation 
of the $next\_timeframe()$ primitive in interface functions.  Recall that a
call to $next\_timeframe()$ is used to advance the HW clock by one cycle.  A call
to the software netlist emulates the effect of one clock period in the RTL 
model. That is, the configuration of state holding elements (registers) and combinational 
elements (wires) are uniform in software netlist and RTL model after every invocation of 
$next\_timeframe()$.  Furthermore, a software
netlist is a sequential model of HW RTL which is represented in C.  Thus, the
composition leads to a single-threaded sequential co-verification model in C.  
%
We now formally define a our co-verification model, denoted by 
$\mathcal{M}_{seq}$.
%
\[ 
\begin{array}[t]{@{}lll}
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{moduleName(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\ 
\end{array}
\]
%
Consider sets of expressions $Exp$ and Boolean expressions $BExp$ 
over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 
$(b \in BExp)$. A $\mathit{CStatement}$ is an assignment,
conditional, sequential concatenation or a loop.
$\mathit{moduleName(Var_1,\dots,Var_n)}$ invokes the top level 
procedure of $\mathcal{SN}$, which is also the top level 
module of the HW RTL, since the translation from RTL to software preserves
the module hierarchy.  The interface variables $\{Var\_1,\dots,Var\_n\}$ 
(input, output and inout ports) are declared inside the 
FW to communicate with $\mathcal{SN}$.  Figure~\ref{compose} shows the 
composition of $\mathcal{SN}$ and FW transaction for UART.
%
\begin{figure}[t]
\scalebox{.50}{\import{figures/}{compose.pspdftex}}
  \caption{Composition of FW and HW transactions}
\label{compose}
\end{figure}

\textbf{Step 4: Property Driven Slicing of Co-verification Model}
A property-driven slicing is performed on the unified model, $\mathcal{M}_{seq}$.  
This step is purely syntactic, meaning that we perform a backward dependency 
analysis starting from the property which only preserve those 
program fragments that are relevant to a given property. The sliced program 
is then passed to the symbolic execution engine for path-based exploration. \\
%
%=======================================================================================
%\subsection{\verifox: Path-wise Symbolic Execution Tool} 
%=======================================================================================
\input{verifox}
%
\subsection{Approach 2: Monolithic Co-verification Technique}
%
\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.6]{figures/traditional_flow.pdf}
\caption{\small Monolithic HW/SW Co-verification Flow in \hwcbmcv}
\label{fig:conventional}
\end{center}
\end{figure}
%
\textbf{Step 1: Co-verification Model}
%
Figure~\ref{fig:conventional} illustrates the monolithic approach to HW/SW
co-veri\-fi\-cation, as implemented in \hwcbmcv.  
%\hwcbmcv is the only open-source
%formal co-verification tool that supports Verilog RTL (System Verilog 2009
%standard) and C/C++ firmware.  
In contrast to path-based approach,
\hwcbmcv maintains two separate flows for HW and FW.  The top flow in
Figure~\ref{fig:conventional} uses synthesis to obtain either a bit-level or
a word-level netlist from Verilog RTL.  A bit-level netlist is represented 
in And Inverter Graph (AIG) format.  Whereas, a word-level netlist is represented 
in format that resembles SMT-LIB2
standards~\footnote{http://smtlib.cs.uiowa.edu/}.  The bottom flow illustrates the
translation of C program into static single assignment (SSA) form. 
Synthesizing netlist from RTL is a standard technique which is widely 
used in hardware model checkers~\cite{DBLP:journals/tse/Clarke76}.
Similarly, translating a C program into a SSA representation is a 
standard technique which is widely used in 
software verification tools, such as CBMC~\cite{cbmc.tacas:2004}. 
\hwcbmcv generates a monolithic co-verification model from AIG and SSA through 
in-tandem symbolic execution of HW and FW components.

\textbf{Step 2: Bounded Model Checking}
%
The co-verification model is checked with bounded model checking (BMC).
Given an unwind depth depth~$k$ and a co-design property $P$, BMC operates
by unwinding the co-verification model~$M$ up to depth~$k$ starting from
initial state $x_0$, represented by an initial state predicate~$I$.
This results in the following formula which is then checked for
satisfiability using an efficient SAT or SMT procedure.
%
\[ I(x_0) \wedge \bigwedge_{i=0}^{k-1} M(x_i, x_{i+1}) \wedge \bigvee_{i=0}^{k} 
\neg P(x_i) \]
%
Thus, BMC exploits the finiteness of $M$ and creates k copies of the
$M$ for each unrolling.

@@

%-------------------------------------------------------------------------------
\subsection{Component level Property}
%-------------------------------------------------------------------------------
A {\em component level} property ($\psi_{c}$) is defined purely at the 
level of individual components of a co-design system. Recall that a component 
may be a HW model, a SW model or an interface model which contains only the 
interface logic.  A component-level property captures the design intent of a 
component in terms of its input/output behaviors.  This type of property 
is useful in a co-design environment for verifying individual components.   
Note that unlike the structure of $\psi_{t}$ which requires the antecedent 
or consequent to be a HW or SW event,  $\psi_{c}$ only contains signals 
(ports) or variables of a particular component.  Figure~\ref{figure:component} 
shows an example of component level property for the UART HW only.  The 
System Verilog Assertion (SVA) is shown in the left and the equivalent 
monitor in C is shown in right.  
%
\begin{figure}[htbp]
\scriptsize
\begin{tabular}{l|l}
\hline
System Verilog Assertions  & Monitor (in C)
\\
\hline
\begin{lstlisting}[mathescape=true,language=Verilog]
P3: assert property 
 (@(posedge clk) 
 ack == 1|-> 
 (valid == 1 && 
 ##2 empty == 0)); 
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
bool monitor_P3() {
 assert(!ack || valid);
 next_timeframe();
 next_timeframe();
 assert(empty==0); }
\end{lstlisting} \\
\hline
\end{tabular}
\caption{Component level property as monitor (in C)}
\label{figure:component}
\end{figure}
%
%-------------------------------------------------------------------------------
\subsection{Debugging Using Counterexample Trace}
%-------------------------------------------------------------------------------
A failure of co-design property implies a fault in HW or a defective FW 
or bug in the interface functions.  Hence, an important feature of a
co-verification tool is its ability to produce a counterexample trace 
in order to localize the bug.  \hwcbmcv generates a counterexample
trace in value change dump (vcd) format that contains concrete assignments to FW
variables, interface signals and HW RTL signals up to a given bound. Whereas, a
counterexample trace in \verifox is a bounded path in the sequential co-verification
model that contains a scenario and its corresponding HW and FW transactions. 
%
%===============================================================================
\section{Experimental Results}
%===============================================================================
%
We report experimental results for FW-HW co-verification of a UART  and a
SoC design.  All our experiments were performed on an Intel Xeon 3.07\,GHz
machine with 48\,GB RAM.  All times reported are in seconds.  The timeout 
for all our experiments was set to 2~hours.  
%

@@

flow in \hwcbmcv. So, we only report bit-level results for \hwcbmcv in this paper. \\ 

@@

Table~\ref{table:safe} shows that path-based symbolic execution dominates
monolithic BMC in all scenarios (marked bold).  The average speedup for
proving safety properties is on average $21\times$ for \verifox over
\hwcbmcv.  The average speedup for detecting bugs is $48\times$ on average 
for \verifox over \hwcbmcv.  Thus, \verifox outperforms \hwcbmcv 
in all scenarios.

@@

\textbf{\emph{About the SoC:}}
%
We obtained an open source System-on-Chip design
from~\cite{DBLP:conf/fmcad/SubramanyanVRM15}.  It consists of an 8051
micro-controller, a memory arbiter, an external memory (XRAM) and
cryptographic accelerators, as shown in Figure~\ref{fig:soc}.  
The design statistics for SoC is given in
Table~\ref{table:stats}.  The accelerator implements encryption/decryption 
using the Advanced Encryption Standard (AES).  There is a separate 
module that interfaces the AES to the 8051 micro-controller using 
a memory-mapped I/O interface.  The micro-controller communicates 
with the accelerators and the XRAM by reading or writing to 
XRAM addresses.  The arbitration of these module is done by 
the memory arbiter module. \\ 
%
\begin{figure}[t]
\scalebox{.50}{\import{figures/}{soc.pspdftex}}
  \caption{SoC Design obtained from~\cite{DBLP:conf/fmcad/SubramanyanVRM15}
\label{fig:soc}}
\end{figure}
%
\textbf{\emph{About the firmware:}}
%
The FW initiates the operation in the SoC by first writing to an
initial memory-mapped register.  The FW implements Linux-style $inb()$
and $outb()$ functions calls, which are used to communicate with the HW
ports.  The FW writes a sequence of non-deterministic data to the XRAM
port and then reads the data from the same port.  The cryptographic
accelerators use direct memory access to fetch the data from the external
memory.  The completion of the operation is determined by polling the
appropriate memory-mapped registers in the FW.
%
\begin{table}
\begin{center}
{
%\begin{scriptsize}
\begin{tabular}{|l|l|l|l|l|}
\hline
  & Bound & \multicolumn{1}{c|}{Monolithic} & \multicolumn{1}{c|}{Path-based} &
  Verification \\ 
\cline{3-4}
  Scenario &  & \multicolumn{1}{c|}{\hwcbmcv} & \verifox & Result \\ 
\cline{3-5}
      &       & Bit-level Netlist & $\mathcal{M}_{seq}$ & Safe/Unsafe \\
\cline{1-5}
      &       &   Time      & Time &  \\
\cline{1-5}      
\multicolumn{5}{|c|}{non-deterministic data and non-deterministic control} \\ \hline
  data\_transfer & 20 & 86.92 & \textbf{17}.\textbf{42} & Safe $(\psi_{t})$ \\
  \hline
  data\_transfer & 30 & 102.92 & \textbf{56}.\textbf{29} & Safe $(\psi_{c})$ \\ 
\hline
\multicolumn{5}{|c|}{non-deterministic data and non-deterministic control} \\ \hline
  data\_transfer & 20 & 92.63 & \textbf{1}.\textbf{27} & Unsafe $(\psi_{c})$ \\ 
\hline
\end{tabular}
%\end{scriptsize}
}
\end{center}
%\vspace{-1.3mm}
\caption{Bounded safety proof times for verification of a SoC 
\label{table:SoC}}
\end{table}
%
\textbf{\emph{Discussion of the result:}}
%
Table~\ref{table:SoC} gives the runtimes for bounded safety proofs of 
co-design properties in SoC.  Column~1 in
Table~\ref{table:SoC} gives the name of the scenario -- transmission with
interrupt (data\_transfer).  Column~2 report the maximum loop bound in the 
FW.  Column~3 present the co-verification times using \hwcbmcv
at the bit-level.  Columns~4 report the verification time using \verifox. 
Column~5 report the verification outcome. \\
\textbf{\emph{Safety properties:}}
%
We verify a total of 7 transaction level propertes and 12 
component level properties for the SoC co-design.  Due to space 
limitations, we describe some representative properties.  We check
whether the acknowledgment for data transmission and data reception arrives
from the micro-controller in the correct cycle.  We verify whether the
non-deterministic data transmitted through $outb()$ is the same
the data received through $inb()$.  We also verify that
reading/writing to the appropriate memory-mapped registers produce the
correct result during the data transmission phase.\\ 
\textbf{\emph{Bug Finding:}} We found one critical \emph{control bug} 
in the SoC design.  The bug is manifested when memory arbiter hardware 
wrongly arbitrates the port selection thereby forcing the write strobe 
for the external RAM to be LOW.  This violates the data transfer protocol 
in the SoC design.

The result in Table~\ref{table:SoC} shows that \verifox is approximately
$5\times$ faster than \hwcbmcv for proving safety.  For detecting bugs, 
the speedup is $72\times$ for \verifox over \hwcbmcv. \\ 
\textbf{\emph{Monolithic versus Path-based:}}
The reason for better scalability of \verifox compared to 
\hwcbmcv is attributed to the path-based exploration strategy combined 
with eager path pruning and incremental SAT solving.  In the case of SoC, 
the FW exercises only the micro-controller and transfer sequence of bytes to the XRAM port 
bypassing peripherals connected to other ports such as hardware accelerator.  
This scenario allows path-based exploration techniques to prune the logic for
accelerator and thereby generate only relevant scenario/transaction pairs.  It 
is important to note that forward symbolic execution without these optimizations 
timeout for all the benchmarks.

>>>>

%
%===============================================================================
\section{Introduction}
%===============================================================================
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                    Pre-RTL versus Post-RTL
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Hardware/software co-verification is broadly performed at two different
phases in the design cycle: pre-Register Transfer Level (pre-RTL) and 
post-RTL.  In the pre-RTL phase, engineers write high-level 
software (SW) models for Intellectual Property, which are commonly 
known as {\em Transaction Level Model}s (TLM)~\cite{codes14}.  The 
TLM models are used to verify high-level functionality of the 
system well before the RTL is available~\cite{codes14,codes15}. 
Co-verification of the TLM and the SW is typically performed 
by means of testing~\cite{hvc,codes14}.

\Omit{
TLMs are typically implemented using the SystemC TLM
library~\cite{tlm} or as plain C programs.  The goal of a TLM is to capture
enough functionality of the hardware (HW) such that it is possible to execute the
SW (drivers or firmware (FW)) against it.  Consequently, the TLMs are
incomplete, as they are designed to contain only those behaviors that are
observable by the SW (the drivers or FW).  In particular, TLMs
are often not cycle-accurate.  These models are used to verify high-level
functionality of the system well before the RTL is
available~\cite{codes14,codes15}. Co-verification of the TLM and the
SW is typically performed by means of testing~\cite{hvc,codes14}.
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                    Challenges in Co-verification 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The verification of a FW given in C/C++ together with an RTL given in 
Verilog/VHDL is a challenging task due to a
few reasons -- 1) the timing and semantic mismatch between the synchronous
clock-driven HW model and asynchronous event-driven SW model. For example, the
FW running on a processor could be much faster than the HW
model it interacts with, 2) lack of support for property specification of 
co-designs that can precisely capture the design intent of the 
communication between FW and HW, 3) large
number of interleavings between HW and FW threads, and 4) lack 
of automated formal tools that support co-designs implemented in C/C++ and 
Verilog/VHDL.   In this paper, we address some of these challenges in 
hardware/software co-verification at post-RTL phase.

@@

\textbf{Step 2: Sequentialization based on Transaction Interaction}
%
Concurrency is at the heart of co-verification.  Testing concurrent 
threads requires exploring all possible interleavings between HW 
and FW threads.  The number of interleavings could potentially be  
exponential.  However, we observe a specific interaction pattern, 
which resembles a producer-consumer relationship.  
That is, a FW thread is mostly independent of the HW 
thread they interact with~\cite{hvc,codes14}.  Specifically, 
a FW thread is only responsible for configuring the 
HW transactions by writing to memory mapped registers, or polling 
the interrupt status register for data transmission, or receiving 
incoming data packets.

@@

Figure~\ref{fig:interleaving1} illustrates the producer-consumer relationship 
in UART . The FW CFG is shown in the left and the HW RTL code snippet 
is shown in the right. We abuse notations and 
represent $\mathcal{HT}$ in the form of RTL and $\mathcal{FT}$ as CFG
for illustration purpose. Here, $\mathcal{HT}$ is the producer of interrupt 
status signal, $irq\_o$, which is read by $\mathcal{FT}$ in state $s_1$. 
The signal $irq\_o$ is  triggered when transmitter buffer is 
{\em EMPTY} or receiver buffer is {\em FULL}. 
$\mathcal{HT}$ also produces interrupt status by writing to the 
memory mapped register, $UART\_IS$, which is consumed by 
$\mathcal{FT}$ in state $s_2$.  In another interaction, $\mathcal{HT}$
sends a data-packet to $UART\_TR$ which is consumed by $\mathcal{FT}$ in state $s_7$.
Similarly, $\mathcal{FT}$ is the producer of $data$, $(outb(data,UART\_TR))$, 
in state $s_6$ which is consumed by $\mathcal{HT}$.  
%
\begin{figure}[t]
\scalebox{.55}{\import{figures/}{interleaving.pspdftex}}
\caption{Producer Consumer Interactions Pattern
\label{fig:interleaving1}}
\end{figure}
%

@@

\textbf{Step 3: Co-verification Model}
%
An unified co-verification model is constructed through appropriate composition of 
FW and its interacting HW transaction.  This is done by instrumenting the 
software netlist model of HW RTL inside the FW transaction at every invocation 
of the $next\_timeframe()$ primitive in interface functions.  Recall that a
call to $next\_timeframe()$ is used to advance the HW clock by one cycle.  A call
to the software netlist emulates the effect of one clock period in the RTL 
model. That is, the configuration of state holding elements (registers) and combinational 
elements (wires) are uniform in software netlist and RTL model after every invocation of 
$next\_timeframe()$.  Furthermore, a software
netlist is a sequential model of HW RTL which is represented in C.  Thus, the
composition leads to a single-threaded sequential co-verification model in C.  
%
We now formally define our co-verification model, denoted by 
$\mathcal{M}_{seq}$.
%
\[ 
\begin{array}[t]{@{}lll}
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{moduleName(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\ 
\end{array}
\]
%
Consider sets of expressions $Exp$ and Boolean expressions $BExp$ 
over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 
$(b \in BExp)$. A $\mathit{CStatement}$ is an assignment,
conditional, sequential concatenation or a loop.
$\mathit{moduleName(Var_1,\dots,Var_n)}$ invokes the top level 
procedure of $\mathcal{SN}$, which is also the top level 
module of the HW RTL, since the translation from RTL to software preserves
the module hierarchy.  The interface variables $\{Var\_1,\dots,Var\_n\}$ 
(input, output and inout ports) are declared inside the 
FW to communicate with $\mathcal{SN}$.  Figure~\ref{compose} shows the 
composition of $\mathcal{SN}$ and FW transaction for UART.
%
\begin{figure}[t]
\scalebox{.50}{\import{figures/}{compose.pspdftex}}
  \caption{Composition of FW and HW transactions}
\label{compose}
\end{figure}

\textbf{Step 4: Property Driven Slicing of Co-verification Model}
A property-driven slicing is performed on the unified model, $\mathcal{M}_{seq}$.  
This step is purely syntactic, meaning that we perform a backward dependency 
analysis starting from the property which only preserve those 
program fragments that are relevant to a given property. The sliced program 
is then passed to the symbolic execution engine for path-based exploration. \\
%
%=======================================================================================
%\subsection{\verifox: Path-wise Symbolic Execution Tool} 
%=======================================================================================
\input{verifox}
%
\subsection{Approach 2: Monolithic Co-verification Technique}
%
\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.6]{figures/traditional_flow.pdf}
\caption{\small Monolithic HW/SW Co-verification Flow in \hwcbmcv}
\label{fig:conventional}
\end{center}
\end{figure}
%
\textbf{Step 1: Co-verification Model}
%
Figure~\ref{fig:conventional} illustrates the monolithic approach to HW/SW
co-veri\-fi\-cation, as implemented in \hwcbmcv.  
%\hwcbmcv is the only open-source
%formal co-verification tool that supports Verilog RTL (System Verilog 2009
%standard) and C/C++ firmware.  
In contrast to path-based approach,
\hwcbmcv maintains two separate flows for HW and FW.  The top flow in
Figure~\ref{fig:conventional} uses synthesis to obtain either a bit-level or
a word-level netlist from Verilog RTL.  A bit-level netlist is represented 
in And Inverter Graph (AIG) format.  Whereas, a word-level netlist is represented 
in format that resembles SMT-LIB2 standards~\footnote{http://smtlib.cs.uiowa.edu/}.  
The bottom flow illustrates the translation of C program 
into static single assignment (SSA) form. 
Synthesizing netlist from RTL is a standard technique which is widely 
used in hardware model checkers~\cite{DBLP:journals/tse/Clarke76}.
Similarly, translating a C program into an SSA representation is a 
standard technique which is widely used in 
software verification tools, such as CBMC~\cite{cbmc.tacas:2004}. 
\hwcbmcv generates a monolithic co-verification model from AIG and SSA through 
in-tandem symbolic execution of HW and FW components.

\textbf{Step 2: Bounded Model Checking}
%
The co-verification model is checked with bounded model checking (BMC).
Given an unwind depth~$k$ and a co-design property $P$, BMC operates
by unwinding the co-verification model~$M$ up to depth~$k$ starting from
initial state $x_0$, represented by an initial state predicate~$I$.
This results in the following formula which is then checked for
satisfiability using an efficient SAT or SMT procedure.
%
\[ I(x_0) \wedge \bigwedge_{i=0}^{k-1} M(x_i, x_{i+1}) \wedge \bigvee_{i=0}^{k} 
\neg P(x_i) \]
%
Thus, BMC exploits the finiteness of $M$ and creates k copies of the
$M$ for each unrolling.

@@

%-------------------------------------------------------------------------------
\subsection{Component level Property}
%-------------------------------------------------------------------------------
A {\em component level} property ($\psi_{c}$) is defined purely at the 
level of individual components of a co-design system. Recall that a component 
may be a HW model, a SW model or an interface model which contains only the 
interface logic.  A component-level property captures the design intent of a 
component in terms of its input/output behaviors.  This type of property 
is useful in a co-design environment for verifying individual components.   
Note that unlike the structure of $\psi_{t}$ which requires the antecedent 
or consequent to be a HW or SW event,  $\psi_{c}$ only contains signals 
(ports) or variables of a particular component.  Figure~\ref{figure:component} 
shows an example of component level property for the UART HW only.  The 
System Verilog Assertion (SVA) is shown on the left and the equivalent 
monitor in C is shown on the right.  
%
\begin{figure}[htbp]
\scriptsize
\begin{tabular}{l|l}
\hline
System Verilog Assertions  & Monitor (in C)
\\
\hline
\begin{lstlisting}[mathescape=true,language=Verilog]
P3: assert property 
 (@(posedge clk) 
 ack == 1|-> 
 (valid == 1 && 
 ##2 empty == 0)); 
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
bool monitor_P3() {
 assert(!ack || valid);
 next_timeframe();
 next_timeframe();
 assert(empty==0); }
\end{lstlisting} \\
\hline
\end{tabular}
\caption{Component level property as monitor (in C)}
\label{figure:component}
\end{figure}
%
%-------------------------------------------------------------------------------
\subsection{Debugging Using Counterexample Trace}
%-------------------------------------------------------------------------------
A failure of co-design property implies a fault in HW or a defective FW 
or bug in the interface functions.  Hence, an important feature of a
co-verification tool is its ability to produce a counterexample trace 
in order to localize the bug.  \hwcbmcv generates a counterexample
trace in value change dump (vcd) format that contains concrete assignments to FW
variables, interface signals and HW RTL signals up to a given bound. Whereas, a
counterexample trace in \verifox is a bounded path in the sequential co-verification
model that contains a scenario and its corresponding HW and FW transactions. 
%
%===============================================================================
\section{Experimental Results}
%===============================================================================
%
We report experimental results for FW-HW co-verification of a UART  and a
SoC design.  All our experiments were performed on an Intel Xeon 3.07\,GHz
machine with 48\,GB RAM.  All times reported are in seconds.  The timeout 
for all our experiments was set to 2~hours.  
%

@@

flow in \hwcbmcv. So, we only report bit-level results for \hwcbmcv in this
paper. \\ 
%\sjcmt{I think this can backfire. The reivewers may ask why were other results not reported if available}\\ 

@@

Table~\ref{table:safe} shows that path-based symbolic execution dominates
monolithic BMC in all scenarios (marked bold).  The average speedup for
proving safety properties is on average $21\times$ 
%\sjcmt{is it average for table 3 only? or also table 4?} 
for \verifox over \hwcbmcv.  The average speedup for detecting bugs is $48\times$ on average 
for \verifox over \hwcbmcv.  Thus, \verifox outperforms \hwcbmcv 
in all scenarios.

@@

\textbf{\emph{About the SoC:}}
%
We obtained an open source System-on-Chip design
from~\cite{DBLP:conf/fmcad/SubramanyanVRM15}.  It consists of an 8051
micro-controller, a memory arbiter, an external memory (XRAM) and
cryptographic accelerators, as shown in Figure~\ref{fig:soc}.  
The design statistics for SoC is given in
Table~\ref{table:stats}.  The accelerator implements encryption/decryption 
using the Advanced Encryption Standard (AES).  There is a separate 
module that interfaces the AES to the 8051 micro-controller using 
a memory-mapped I/O interface.  The micro-controller communicates 
with the accelerators and the XRAM by reading or writing to 
XRAM addresses.  The arbitration of these module is done by 
the memory arbiter module. \\ 
%
\begin{figure}[t]
\scalebox{.50}{\import{figures/}{soc.pspdftex}}
  \caption{SoC Design obtained from~\cite{DBLP:conf/fmcad/SubramanyanVRM15}
\label{fig:soc}}
\end{figure}
%
\textbf{\emph{About the firmware:}}
%
The FW initiates the operation in the SoC by first writing to an
initial memory-mapped register.  The FW implements Linux-style $inb()$
and $outb()$ functions calls, which are used to communicate with the HW
ports.  The FW writes a sequence of non-deterministic data to the XRAM
port and then reads the data from the same port.  The cryptographic
accelerators use direct memory access to fetch the data from the external
memory.  The completion of the operation is determined by polling the
appropriate memory-mapped registers in the FW.
%
\begin{table}
\begin{center}
{
\begin{scriptsize}
\begin{tabular}{|l|l|l|l|l|}
\hline
  & Bound & \multicolumn{1}{c|}{Monolithic} & \multicolumn{1}{c|}{Path-based} &
  Verification \\ 
\cline{3-4}
  Scenario &  & \multicolumn{1}{c|}{\hwcbmcv} & \verifox & Result \\ 
\cline{3-5}
      &       & Bit-level Netlist & $\mathcal{M}_{seq}$ & Safe/Unsafe \\
\cline{1-5}
      &       &   Time      & Time &  \\
\cline{1-5}      
\multicolumn{5}{|c|}{non-deterministic data and non-deterministic control} \\ \hline
  data\_transfer & 20 & 86.92 & \textbf{17}.\textbf{42} & Safe $(\psi_{t})$ \\
  \hline
  data\_transfer & 30 & 102.92 & \textbf{56}.\textbf{29} & Safe $(\psi_{c})$ \\ 
\hline
\multicolumn{5}{|c|}{non-deterministic data and non-deterministic control} \\ \hline
  data\_transfer & 20 & 92.63 & \textbf{1}.\textbf{27} & Unsafe $(\psi_{c})$ \\ 
\hline
\end{tabular}
\end{scriptsize}
}
\end{center}
%\vspace{-1.3mm}
\caption{Bounded safety proof times for verification of a SoC 
\label{table:SoC}}
\end{table}
%
\textbf{\emph{Discussion of the result:}}
%
Table~\ref{table:SoC} gives the runtimes for bounded safety proofs of 
co-design properties in SoC.  Column~1 in
Table~\ref{table:SoC} gives the name of the scenario -- transmission with
interrupt (data\_transfer).  Column~2 report the maximum loop bound in the 
FW.  Column~3 present the co-verification times using \hwcbmcv
at the bit-level.  Columns~4 report the verification time using \verifox. 
Column~5 report the verification outcome. \\
\textbf{\emph{Safety properties:}}
%
We verify a total of 7 transaction level propertes and 12 
component level properties for the SoC co-design.  Due to space 
limitations, we describe some representative properties.  We check
whether the acknowledgment for data transmission and data reception arrives
from the micro-controller in the correct cycle.  We verify whether the
non-deterministic data transmitted through $outb()$ is the same as 
the data received through $inb()$.  We also verify that
reading/writing to the appropriate memory-mapped registers produce the
correct result during the data transmission phase.\\ 
\textbf{\emph{Bug Finding:}} We found one critical \emph{control bug} 
in the SoC design.  The bug is manifested when memory arbiter hardware 
wrongly arbitrates the port selection thereby forcing the write strobe 
for the external RAM to be LOW.  This violates the data transfer protocol 
in the SoC design.

The result in Table~\ref{table:SoC} shows that \verifox is approximately
$5\times$ faster than \hwcbmcv for proving safety.  For detecting bugs, 
the speedup is $72\times$ for \verifox over \hwcbmcv. \\ 
\textbf{\emph{Monolithic versus Path-based:}}
The reason for better scalability of \verifox compared to 
\hwcbmcv is attributed to the path-based exploration of a 
sequential co-verification model using eager path pruning and 
incremental SAT solving to automatically mine relevant scenario/transaction
pairs.  In the case of SoC, the FW exercises only the micro-controller 
and transfer sequence of bytes to the XRAM port bypassing peripherals 
connected to other ports such as hardware accelerator.  
This scenario allows path-based exploration techniques to prune the logic for
accelerator and thereby generate only relevant scenario/transaction pairs.  It 
is important to note that forward symbolic execution without these optimizations 
timeout for all the benchmarks.
